---
title: "Treating the data"
format: html
editor: visual
---

After better understanding our dataset, we are now ready to prepare the data for our forecasting exercise. We have already treated part of the data in the previous script: the sales information was aggregated to category/week, our hierarchical level of interest. This script will then focus on treating `calendar` and `prices` and joining all variables in a single file that will be used in our forecasting comparison.

## Packages

The packages used in this document are:

```{r}
#| warning: false
#| message: false
options(repos = c(CRAN = "https://cran.r-project.org"))

install.packages("tidyverse")
install.packages("magrittr")

library(tidyverse)
library(magrittr)

```

## Loading the data

```{r}
#| warning: false
#| message: false

# Loading prices
prices<-read_csv("../data/sell_prices.csv")

# Weights - will be used with prices
w<-read_csv("../data/weights_validation.csv")

```

## Prices

```{r}
#| warning: false
#| message: false

head(prices)

```

We begin by aggregating the product/week prices across all stores as in the last script:

Note: we are considering "category level (cat_id)". The code needs to be adapted for other levels.

```{r}
#| warning: false
#| message: false

prices$cat_id<-unlist(str_split(prices$item_id, "_", 2, simplify = T), recursive = F)[,1]
w %<>% filter(Level_id == "Level12") %>%
  select(item_id = Agg_Level_1, store_id = Agg_Level_2, dollar_sales = Dollar_Sales)
prices %<>% left_join(w, by = c("item_id", "store_id"))
prices %<>% group_by(cat_id, item_id, wm_yr_wk) %>%
  summarise(sell_price = weighted.mean(sell_price, dollar_sales),
            dollar_sales = sum(dollar_sales)) %>%
  ungroup()
prices$sell_price[prices$dollar_sales == 0]<-0

head(prices)

```

```{r}
#| warning: false
#| message: false
#| echo: false

rm(w)

```

Below we create the weighted mean, median and the 25%, 75% and 90% quantiles of weekly prices for each `cat_id`. The usage of more upper quantiles than lower ones is due to the observed right-skewed distributions in two out of three categories.

```{r}
#| warning: false
#| message: false

# Summarizing
prices %<>% group_by(cat_id, wm_yr_wk) %>%
  summarise(price_wmean = weighted.mean(sell_price, dollar_sales),
            price_median = median(sell_price),
            price_q25 = quantile(sell_price, 0.25),
            price_q75 = quantile(sell_price, 0.75),
            price_q90 = quantile(sell_price, 0.9)) %>%
  ungroup()

head(prices)

write_csv(prices, "../data/prices.csv")
```
